---
title: "Board Report Parser"
author: "Ann Scruggs"
date: "2024-12-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Transforming the Dataset

### Setup

The code chunk below loads the appropriate data downloaded from PowerBI and does some initial setup

```{r part_1}

######################################################################
#Loading in Data and initial setup
######################################################################

#install.packages("readxl") 
library(readxl)
#install.packages("dplyr")
library(dplyr)

#setting directory and loading data
setwd("~/Documents/Serge/Metrics/Dec_BR/")
data <- read_excel("data.xlsx")

#Cleaning data
field_data <- data %>%
  mutate(Category = ifelse(Category == "Renewal", "Renew", Category)) %>% #Modifying "Renewal" category name
  filter(Name...1 != "Total") %>% #removing total row
  filter(Category != "Team Life") #filtering out TeamLife rows 
```

### Sub-Category Assignment

The code chunk below creates a new column that assigns a sub - category to each activity. New and unassigned activities can be added to this chunk as they come.

```{r part_2}

######################################################################
#Adding "sub_category" column and assigning each activity
######################################################################

categorized_data <- field_data %>%
  mutate(sub_category = case_when(
    
    #REACH SUB-CATEGORIES
    ##Pre-E Encounter
    Name...2 %in% c("Intentional relational encounter with unbeliever", 
                    "People participating in outreach events", 
                    "Pre-E Encounters", 
                    "Number of hours engaged in relationship-building ministry", 
                    "New NA contacts added to the team's relational network through a face-to-face en",
                    "Intentional relational encounter with unbeliever from ministry focus group",
                    "Intentional relational encounter with unbeliever not from ministry focus group",
                    "Number of those who received practical assistance") ~ "Pre E Encounter",
    ##Gospel Conversation
    Name...2 %in% c("Number of patient CDs as reported by hospital chaplains",
                    "G-spl conversations",
                    "Spiritual Conversations with Non-Believers",
                    "Special Outreach Events (EV emphasis)",
                    "Number of significant new EV encounters with NBs",
                    "Conversations that led to speaking about the gspl",
                    "E Encounters",
                    "Significant EV conversations",
                    "Number of EV encounters",
                    "Number of ABs",
                    "Number of pre-CD conversations by team with non believers",
                    "Number of times a team member talked about God with non-believer(s)",
                    "Verbal gospel encounter with unbeliever from ministry focus group",
                    "Verbal gospel encounter with unbeliever not from ministry focus group") ~ "Gospel Conversation",
    ##Formal Gospel Encounter
    Name...2 %in% c("Number of times a team member preached or taught",
                    "Event attendance where team member taught of preached",
                    "Number of times a team member preached or provided spiritual teaching",
                    "Students taught in English & storytelling classes",
                    "Worship",
                    "Special Outreach Events (EV emphasis)",
                    "Number of events at which team members taught or preached",                     
                    "Number of times team member preached/taught in small group",
                    "Number of people attending HC, C or BS - NB or B",
                    "Number of other classes taught",
                    "Large group teaching",
                    "Sunday church attendances: total unbelievers") ~ "Formal Gospel Encounter",
    
    
    #RENEW SUB-CATEGORIES
    ##Discipleship Encounter
    Name...2 %in% c("People receiving mentoring/discipling",
                    "Number of people our team is discipling",
                    "Number of people engaged in small group Book studies by team members",
                    "People receiving spiritual/marriage coaching",
                    "Discipleship encounter",
                    "Meaningful encounters with women involved in bakery",
                    "Number of People Team is discipling",
                    "Number people engaged in mentoring and discipling relationships",   
                    "Number of discipleship encounters per month",
                    "One to one (inward)",                                                             
                    "One to one (outward)",
                    "Relational prayer times",                                                         
                    "Small group teaching",
                    "D encounters", 
                    "Number of Mentoring/Discipling meetings with NA people",
                    "Number of people team members discipled or mentored",
                    "Number of times a team member was involved in MOP",                               
                    "Number of interns/apprentices receiving discipleship/mentoring",                  
                    "Mentoring and discipling meetings",                                               
                    "Group Bible study attendances") ~ "Discipleship Encounter",
    ##Train & Support local leaders
    Name...2 %in% c("Number of med students & interns rotating through the hospital per month",
                    "Number of local leaders/ministries encouraged",
                    "Number of NA people attending IRG",
                    "Number of NA people attending Oasis Group",
                    "Days per month team member provided training for national nurses/nursing student",
                    "Number of med students & medical interns receiving clinical training",
                    "Colleague Collaboration",
                    "Partnership meetings",
                    "Small group teaching",
                    "Number of missionaries/expats who received mentoring or prayer",                 
                    "Number of national leaders who received mentoring or prayer",                   
                    "Number of national leaders' wives who received mentoring or MOP",
                    "Number of med students and interns rotating through the hospital per month" ,
                    "Number of people engaged in large group PT/training events.",
                    "People trained at events",
                    "Number of people given instruction or training",
                    "Number of times team members met with local ministry partners/project leaders",
                    "Training event attendances to develop local ministry leaders",
                    "People attending medical seminars",
                    "Encounters encouraging local leaders / ministries") ~ "Train & Support local leaders",
    ##Church Plant Activity
    Name...2 %in% c("Number of people engaged in small group Book studies by team members",
                    "Participation in local church activities (not Sunday worship)",
                    "Number of HCes meeting",
                    "Sunday church attendances",
                    "Corporate prayer meetings") ~ "Church Plant Activity",
    
    
    #RESTORE SUB-CATEGORIES 
    ##Medical Care
    Name...2 %in% c("# of patients seen in outpatient consultation (including by Burundian doctors)",
                    "Number of patients hospitalized at Kibuye Hope Hospital",
                    "Patients served by the Kibuye Hope Hospital Needy Patient Fund",
                    "Total number of patient encounters per month",
                    "Patients seen in clinic",
                    "Number of clinic visits in zone 3 for diabetes or general urgent care",
                    "Number of deliveries at Kapsowar Hospital",                                       
                    "Number of med/peds inpatients seen at Kapsowar Hospital",  
                    "Number of patients seen in outpatient consultation by RIIO staff/residents",
                    "Number of surgeries performed by RIIO staff/residents or Eye Love Africa outreac",
                    "Number of patients seen in outpatient consultation") ~ "Medical Care",
    ##Mercy, diaconal, justice ministry
    Name...2 %in% c("Number of jobs provided",
                    "Number of those encountering mercy and justice",
                    "Activity related to Crucero Bakery Pilot Project",                               
                    "Encounters encouraging local leaders / ministries",                              
                    "Event as a result of empowering locals",
                    "Number of vulnerable people served food or clothing",                            
                    "Individual mercy/diaconal encounter",                                            
                    "One-off group activity attendances",                                             
                    "Recurring group activity attendances") ~ "Mercy, diaconal, justice ministry",
    ##Counseling Sessions
    Name...2 %in% c("Number of counseling sessions provided", 
                    "Number of people given counseling") ~ "Counseling Sessions",
    
    
    ##Renewal Team
    Name...1 == "Renewal" ~ "Renewal Team",
    
    #UNSASSIGNED ACTIVITIES
    
    TRUE ~ "Unassigned"))

head(categorized_data)
```

### Troubleshooting

The code chunk below is good for troubleshooting.

```{r part_3}

######################################################################
#Adding Notes column to explain why some unassigned activities should not be included in the sub_category count
######################################################################

categorized_data <- categorized_data %>%
  mutate(Notes = case_when(
    
    #Double-Counts
    Name...2 == "# of surgeries performed at KHH (inc. C-sections & procedures by Burundian Drs)"
    ~ "already counted with 'Number of patients hospitalized at Kibuye Hope Hospital'",
    
    Name...2 %in% c("Number of Internally Displaced Persons provided with medical care", 
                    "Number of surgeries performed") ~ "already counted with 'Total number of patient encounters per month'",
    
    Name...2 %in% c("Functional church members serving in regular role or ministry of church",         
                    "Sunday church attendance: persons from ministry focus group",                    
                    "Sunday church attendance: workers sent by overseas orgs and their minor children",
                    "Functional church members") ~ "already counted with 'Sunday church attendances'",
    
    Name...2 == "Recurring group activity attendances: persons from ministry focus group"
    ~ "already counted with 'Recurring group activity attendances'",
    
    
    #Team, Donor, or HomeOffice Related
    Name...2 %in% c("Meetings designed for team care",
                    "Personal retreat days",
                    "Facilitating short-term partnerships",
                    "Number of supporting churches/donors engaged with",
                    "Members attended Sunday service this month",
                    "Number of adult team members who took a personal retreat day",
                    "Number of supporters contacted individually",
                    "Team Meetings",                                                                   
                    "Number of retreat days per month",                                               
                    "Cultural learning experiences. Please describe.",                                 
                    "J Days",                                          
                    "Recruiting to the SA team or summer internship",   
                    "Number of fundraising meetings",                                                  
                    "Number of hours spent in language or cultural learning activities",               
                    "Number of hours spent in mass communication",
                    "Number of meetings/gatherings designed for spiritual/emotional team care",
                    "Number of visitors/volunteers hosted", 
                    "Number of personal retreats or prayer days",                                     
                    "Number of visitors hosted",                                                      
                    "Number of social events (official and unofficial) with team members from other t",
                    "Personal Days Away and Prayer Times",                                             
                    "Language learning hours ",
                    "P times",                                                                        
                    "Number of hours spent in personal or team retreat days", 
                    "Number of supporters/supporting churches with whom team members skyped/zoomed",   
                    "Number of substantial intra team activities",     
                    "Spiritual or emotional care encounter",
                    "Number of hours team members contributed to Srg or field training materials pr",
                    "Professional development encounter") ~ "Team, Donor, or HomeOffice Related Activiy"))


######################################################################
#filters for all categorized_data that has an unassigned sub_category with no notes - these are the activities to take another look at (should they be slotted in a category?)
######################################################################

unassigned_data <- categorized_data %>%
  filter(sub_category == "Unassigned" & is.na(Notes))

unique(unassigned_data$Name...2)

```

# Overseas Visualizations

Creating the dataset to use in the plots and setting brand colors

```{r part_4}
#install.packages("plotly")
library(plotly)

overseas_data <- categorized_data %>%
  filter(!sub_category %in% c("Unassigned", "Renewal Team")) #filtering out any "unassigned" rows

brand <- c("#EF5A4A","#55373C","#0E8BA4","#F5BA40","#AC4432", "#443F32","#388166", "#0B473A")

rrr_brand <- c("Restore" = "#EF5A4A", "Reach" = "#388166","Renew" = "#0E8BA4")
```

### Total RRR Plot

```{r combined_plot}

rrr_plot <- plot_ly(overseas_data, labels = ~Category, values = ~Quantity, type = 'pie', marker = list(colors = brand),
        showlegend = FALSE, textinfo = 'label+value')

rrr_plot <- rrr_plot %>% layout(
  paper_bgcolor = "#EFECE0",
  font = list(
    family = "Open Sans",        # Set font family (optional)
    size = 28,              # Set font size for labels
    color = "white"       # Set font color for labels
  ))

rrr_plot 

#to save interactive plot 
#htmlwidgets::saveWidget(rrr_plot, "~/Documents/Serge/Metrics/Dec_BR/Leadership-Reports/RRR_plot.html")
```

### Total Activity Plot

```{r total_activity_plot}

color_repeated <- rep(brand, length.out = 30)

total_activity_plot <- plot_ly(overseas_data, labels = ~Name...2, values = ~Quantity, 
        type = 'pie', textinfo ='percent', 
        showlegend = FALSE, marker = list(colors = color_repeated),
        insidetextorientation = 'radial',  # Make the inside text go radial (good for fitting)
        textposition = 'inside')  # Place text inside each slice

total_activity_plot <- total_activity_plot %>% layout(
  paper_bgcolor = "#EFECE0")

total_activity_plot

#to save interactive plot 
#htmlwidgets::saveWidget(total_activity_plot, "~/Documents/Serge/Metrics/Dec_BR/Leadership-Reports/total_activity_plot.html")
```

### Reach Plot

```{r reach_plot}

#REACH data
reach_data <- overseas_data %>%
  filter(sub_category %in% c("Pre E Encounter","Gospel Conversation","Formal Gospel Encounter"))

#Plot
reach_plot <- plot_ly(reach_data, labels = ~sub_category, values = ~Quantity, type = 'pie', 
                      textinfo = 'label+percent', marker = list(colors = brand), showlegend = FALSE,
                      insidetextorientation = 'horizontal', textposition = 'inside')

reach_plot <- reach_plot %>% layout(
  font = list(
    family = "Open Sans",        # Set font family (optional)
    size = 28,              # Set font size for labels
    color = "white"       # Set font color for labels
  ))

reach_plot

#to save interactive plot 
#htmlwidgets::saveWidget(reach_plot, "~/Documents/Serge/Metrics/Dec_BR/Leadership-Reports/reach_plot.html")
```

### Renew Plot

```{r renew_plot}

#RENEW data
renew_data <- overseas_data %>%
  filter(sub_category %in% c("Discipleship Encounter","Train & Support local leaders","Church Plant Activity"))

#Plot
renew_plot <- plot_ly(renew_data, labels = ~sub_category, values = ~Quantity, type = 'pie', 
                      textinfo = 'label+percent', marker = list(colors = brand), showlegend = FALSE,
                      insidetextorientation = 'horizontal', textposition = 'inside')
      
renew_plot <- renew_plot %>% layout(
  font = list(
    family = "Open Sans",        # Set font family 
    size = 28,              # Set font size for labels
    color = "white"       # Set font color for labels
  ))

renew_plot

#to save interactive plot 
#htmlwidgets::saveWidget(renew_plot, "~/Documents/Serge/Metrics/Dec_BR/Leadership-Reports/renew_plot.html")
```

### Restore Plot

```{r restore_plot}

#RESTORE data
restore_data <- overseas_data %>%
  filter(sub_category %in% c("Medical Care","Mercy, diaconal, justice ministry","Counseling Sessions"))

#Plot
restore_plot <- plot_ly(restore_data, labels = ~sub_category, values = ~Quantity, type = 'pie', textinfo = 'label+percent', 
                     marker = list(colors = c("#388166","#0E8BA4")), showlegend = FALSE,insidetextorientation = 'horizontal')
                     

restore_plot <- restore_plot %>% layout(
  font = list(
    family = "Open Sans",        # Set font family (optional)
    size = 28,              # Set font size for labels
    color = "black"),
  margin = list(t = 110, b = 15, l = 20, r =10)# Set font color for labels
)

restore_plot

#to save interactive plot 
#htmlwidgets::saveWidget(restore_plot, "~/Documents/Serge/Metrics/Dec_BR/Leadership-Reports/restore_plot.html")
```

# Specific Outputs

The code chunk below shows the values for each sub_category and is used to accurately fill in the side comments to the leadership report slides

```{r specific outputs}

#REACH comments
reach_values <- reach_data %>%
  group_by(sub_category) %>%
  summarise(sum(Quantity))

#RENEW comments
renew_values <- renew_data %>%
  group_by(sub_category) %>%
  summarise(sum(Quantity))

#RESTORE comments
restore_values <- restore_data %>%
  group_by(sub_category) %>%
  summarise(sum(Quantity))
```

# Renewal Team Visualizations

```{r Renewal}

total_data <- categorized_data %>%
  filter(sub_category != "Unassigned") #filtering out any "unassigned" rows

total_rrr_plot <- plot_ly(total_data, labels = ~Category, values = ~Quantity, type = 'pie', 
                          marker = list(colors = brand), showlegend = FALSE, textinfo = 'label+value',
                          insidetextorientation = 'horizontal', textposition = 'inside')

total_rrr_plot <- total_rrr_plot %>% layout(
  paper_bgcolor = "#EFECE0",
  font = list(
    family = "Open Sans",        # Set font family (optional)
    size = 20            # Set font size for labels
  ))

total_rrr_plot 
```

```{r Renewal only}

renewal_data <- total_data %>%
  filter(sub_category == "Renewal Team")


renewal_plot <- plot_ly(renewal_data, labels = ~Category, values = ~Quantity, type = 'pie', 
                          marker = list(colors = "#388166"), showlegend = FALSE, textinfo = 'label+value',
                          insidetextorientation = 'horizontal', textposition = 'inside')

renewal_plot <- renewal_plot %>% layout(
  paper_bgcolor = "#EFECE0",
  font = list(
    family = "Open Sans",        # Set font family (optional)
    size = 20            # Set font size for labels
  ))

renewal_plot
```
